Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-26 files/Chapter3/femconvergence.tex
chapter 3
Chapter 3

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{FEM computation of eigenvalues and eigenfunctions}
Let $H$ be a Hilbert space with the inner product $(\cdot,\cdot)$ and induced norm $||\cdot||$. Let $V$ be a linear subspace of $H$, with inner product defined by the bilinear form $b(\cdot,\cdot)$. It is assumed that this bilinear form $b$ is symmetric.

\subsection{General eigenvalue problem}

The following eigenvalue problem is considered in \cite{SF73}.

@@

Suppose $E^h$ is a finite dimensional subspace of $V$. Consider the Galerkin approximation for \eqref{GEVP}:
\subsubsection*{Problem $E^h$}
Find $u^h \in S^h$ such that $u^h \neq 0$ and \[b(u^h, v) = \lambda^h(u^h,v) \textrm{ for all } v \in S^h.\]

For detail, see Chapter 5.

\textcolor{red}{********** need to move this ******************}

\subsection{Preliminary theory and notation}
In this section, the work in the textbook \cite{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for greater clarity.

\textcolor{red}{********** need to move this ******************}

\subsubsection*{Properties of eigenvalues}
It is well known that the eigenvalues of a eigenvalue problem such as \eqref{GEVP} form an infinite sequence of eigenvalues $(\lambda_i)$. Assume that the eigenvectors can be ordered in such a way that $\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...$ where $\lambda_i$ are the corresponding eigenvalues.

Since any multiple of a eigenfunction is still an eigenfunction, the eigenfunctions can be normalized so that $||u_i|| = 1$ for all $i$.

Problem $E^h$ can be written as a matrix eigenvalue problem,
\begin{eqnarray*}
	\lambda^nM\bar{u}_n = K\bar{u}_n.
\end{eqnarray*} The pair $(\lambda^n, \bar{u}_n)$ correspond to the pair $(\lambda^h_k, u^h_k)$ which is the solution of Problem $E^h$.

In a finite-dimensional subspace, the ordering of vectors is preserved as it is in the original space. Denote the normalized eigenvectors in the space $S^h$ as $u^h_k$ with corresponding eigenvalues $\lambda^h_i$ for $k = 1,2,...,m$.

Also any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}

\subsubsection*{Properties of eigenvalues}
It is well known that the eigenvalues of a eigenvalue problem such as \eqref{GEVP} form an infinite sequence $(\lambda_i)$. Assume that the eigenvalues $\lambda_i$ can be ordered $\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...$ and denote the corresponding eigenfunctions $u_i$.

Since any multiple of a eigenfunction is still an eigenfunction, the eigenfunctions can be normalized so that $||u_i|| = 1$ for all $i$.

@@

Also any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}

\subsection{Estimating the eigenvalues.} \label{sec:est}
\textbf{Rayleigh quotient}
\begin{equation}
	R(v) = \frac{b(v,v)}{c(v,v)} \ \text { for } v \in V. \label{Rayleigh}
\end{equation} \label{sym:Rayleigh}

\textbf{minmax Principle}\\
Let T denote the class of subspaces of $V$ that jas dimension $j$. Then
\begin{equation*}
	\lambda_j = \min_{s\in T}\max_{v \in S} R(v).
\end{equation*}

\textbf{Rayleigh-Ritz Projection}\\
If $u \in V$, then $Pu$ is its projection in the subspace $S^h$.
\begin{equation*}
	b(u-Pu,v^h) = 0 \ \text{ for all } v \in S^h.
\end{equation*}

@@

For the requirements of this dissertation, this can be simplified with an assumption. The assumption is that the basis of $S^h$ consists of polynomials. With this assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}|| \leq C h^{k-m} || u^{(k)}|| \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

This will be referred to as the approximation theorem in this section.

Let $E_j \in V$ denote the eigenspace spanned by the exact eigenvectors $\left\{u_1,u_2,...,u_j \right\}$ for $j = 1,2,...,m$.

Then consider the subspace $S_j$ of $S^h$ where
\begin{equation*}
	S_j = PE_j \ \text{ for } j = 1,2,...,m.
\end{equation*} The elements $Py_i$ are the projections of the eigenfunctions $u_i$ into the space $S^h$. These projections $Py_i$ are not necessarily equal to $u^h_j \in S^h$, but it can be close. The situation is not simple. It is possible that $S_j$ is not a subspace of $S^h$. Assume that the dimension of $S^h$ is large enough

\vspace{20mm}

(It is possible that $Py_k = 0$ for large $k$.)

Let $B_m = \left\{u \in e_m \ | \ ||u|| = 1 \right\}$ and define $\mu_m = \inf\left\{(Pu,Pu \ | \ u \in B_m)\right\}$.

The first step to obtain estimates for the eigenvalues, it to show that the elements of $B_m$ are linearly independent. In the first part we follow the approach in \cite{Zie2000} (A Computable Spectral Theorem - Martin Ziegler - 2000 ?). The author introduced the quantity $\mu_m$.

\newtheorem{Prop_2}[Prop_1]{Proposition}
\begin{Prop_2}
	\label{Prop_2}
	$\lambda^{h}_{m} \leq \max R(Pu) \ \text{ for } u \in B_{m}$
\end{Prop_2}
\begin{proof}
	Since $\dim S_{m} = m$, following from the minmax principle
	\begin{equation*}
		\lambda_m^h \leq \max R(v) \ \text{ for } v \in S_m.
	\end{equation*}

	Now we take an arbitrary $v \in S_{m}$, $v \neq 0$. Then there exists a $u \in E_{m}$ such that $Py = v$. This $Py$ is the projection into $S_m$ of some $u \in E_m$ (which is also $\displaystyle \frac{1}{||u||}u \in B_m$).

	Next we show that $R(||u||^{-1}u) = R(u)$:

	\[ \frac{b\left(\frac{1}{||u||}u,\frac{1}{||u||}u\right)}{c\left(\frac{1}{||u||}u,\frac{1}{||u||}u\right)} = \frac{b(u,u)}{c(u,u)} = R(u). \]

	So finally,
	\begin{eqnarray*}
	\lambda_{m}^{h} &\leq & \max R(v) \ \text{ for } \ v \in S_{m},\\
						&=& \max R(Pu) \ \text{ for } \ u \in E_{m},\\
						&=& \max R(Pu) \ \text{ for } \ u \in B_{m}.
	\end{eqnarray*}
\end{proof}

Pick any $i,j \in \mathbb{N}$ such that $i,j \leq m$. Let $v_{i},v_{j} \in V$ be arbitrary, then
\begin{eqnarray*}
	b(u_{i},u_{i}) &=& \lambda_{i}c(u_{i},u_{i}),\\
	b(u_{j},u_{j}) &=& \lambda_{j}c(u_{j},u_{j}).
\end{eqnarray*}
Now pick $u_{i} = u_{j}$ and $u_{j} = u_{i}$. Then
\begin{eqnarray*}
	b(u_{i},u_{j}) &=& \lambda_{i}c(u_{i},u_{j}),\\
	b(u_{j},u_{i}) &=& \lambda_{j}c(u_{j},u_{i}).
\end{eqnarray*}
Then using the symmetry of $b(\cdot,\cdot)$ and $c( \cdot, \cdot )$, and subtracting,
\begin{eqnarray*}
	0 &=& (\lambda_{i} - \lambda_{j})c( u_{i}, u_{j} )
\end{eqnarray*}
So if $i \neq j$ then $\lambda_{i} \neq \lambda_{j}$. Therefore $u_{i}$ and $u_{j}$ must be orthogonal.

And as mentioned previously, $||u|| = 1$.

	The summation parameters can be merged into a single parameter. Then
	\begin{eqnarray*}
		b(u,u)  & = & \sum_{i=1}^{m} c_{i}^{2} \lambda_{i} u_i,\\
				& \leq & \lambda_{m}\sum_{i=1}^{m} c_{i}^{2} u_i,\\
				& = & \lambda_{m}||u||^2.
	\end{eqnarray*} for all $u \in B_m$.\\

	Using the Rayleigh quotient, and the definition of $\mu_{m}^{h}$, 
	\begin{eqnarray*}
		R(Pu) &=& \frac{b(Pu,Pu)}{c(Pu,Pu)},\\
			&=& \frac{b(Pu,Pu)}{||Pu||^{2}},\\
			&\leq & \frac{\lambda_{m}}{\mu_{m}^{h}}.
	\end{eqnarray*}

Following Lemma \ref{Lem_1}, and since $\lambda_{i}^{h} \geq \lambda_{i}$ it follows that $0 < \mu_{m}^{h} \leq 1$. It is now possible to define the `error bound' in \cite{SF73}:
\begin{eqnarray*}
	\sigma_{m}^{h} = 1 - \mu_{m}^{h}.
\end{eqnarray*}

\newtheorem{Cor_1}[Prop_1]{Proposition}
\begin{Cor_1}
	$0 \leq \sigma_{m}^{h} < 1$ and $\lambda_{m}^{h} - \lambda_{m} \leq \lambda_{m}^{h}\sigma_{m}^{h}$
\end{Cor_1}
\begin{proof}
	Starting with the result of Lemma \ref{Lem_1},
	\begin{align*}
		\lambda_{m}^{h} &\leq \frac{\lambda_{m}}{\mu_{m}^{h}}, \\
		\lambda_{m}^{h}\mu_{m}^{h} &\leq \lambda_{m}. \\
	\end{align*}
	This is equivalent to
	\begin{align*}
		-\lambda_{m} &\leq -\lambda_{m}^{h}\mu_{m}^{h}. 
	\end{align*}
	Add $\lambda_{m}^{h}$ to both sides
	\begin{align*}
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h} - \lambda_{m}^{h}\mu_{m}^{h}. 
	\end{align*}
	And finally using the error bound $\sigma_{m}^{h} = 1 - \mu_{m}^{h}$,
	\begin{align*}
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h}\sigma_{m}^{h}.
	\end{align*}
\end{proof}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{FEM computation of eigenvalues and eigenfunctions}
\textcolor{red}{Change H to W?}

Let $H$ be a Hilbert space with the inner product $c(\cdot,\cdot)$ and induced norm $||\cdot||_{W}$. Let $V$ be a linear subspace of $H$, with inner product defined by the bilinear form $b(\cdot,\cdot)$. It is assumed that the bilinear form $b$ is symmetric and that the assumptions in Section \ref{sec:existence:ModalAnalysis} holds.

The following eigenvalue problem is considered in \cite{SF73}. The same problem was treated in Section \ref{sec:existence:ModalAnalysis}, although the notation differs slightly.

@@

Recall that the eigenvectors can be ordered in such a way that \[\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...\] where $\lambda_i$ are the corresponding eigenvalues.

\subsubsection*{Properties of eigenvalues}
Since any multiple of a eigenfunction is still an eigenfunction, the eigenfunctions can be normalized so that $||u_i||_{W} = 1$ for all $i$.

Let $\left\{ \phi_k \in V \ | \  k = 1,2,...,N_e \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_k \in V \ | \ k = 1,2,...,N_e\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

Consider the Galerkin approximation for \eqref{GEVP}:
\subsubsection*{Problem $E^h$}
Find $u^h \in S^h$ such that $u^h \neq 0$ and \[b(u^h, v) = \lambda^h(u^h,v) \textrm{ for all } v \in S^h.\]

For examples, see Chapter 5.

Problem $E^h$ can be written as a matrix eigenvalue problem,
\begin{eqnarray}
	\lambda^nM\bar{u}_n = K\bar{u}_n. \label{eq:matrix_eigenvalue}
\end{eqnarray}

Since $N_e$ is never small and usually large to very large, a compute algorithm is required to calculate the eigenvalues (and eigenfunctions) of \eqref{eq:matrix_eigenvalue}.

The pair $(\lambda^n, \bar{u}_n)$ correspond to the pair $(\lambda^h_k, u^h_k)$ which is the solution of Problem $E^h$. It is necessary to some of the teory to make the connection.

In $S^h$, the ordering of vectors is the same as in the original space. Denote the normalized eigenvectors in the space $S^h$ as $u^h_k$ with corresponding eigenvalues $\lambda^h_k$ for $k = 1,2,...,N_e$.

In section \ref{sec:estimating_the_eigenvalues} it is proved that the error $| \lambda_k^h - \lambda_k|$ is large when $k$ is large. It is for example possible that $|\lambda^h_1 - \lambda_1|$ is sufficiently small while $\lambda^h_k$ cannot even be considered an approximation of $\lambda_k$ when $\displaystyle k > \frac{1}{2}N_e$.

Finally, any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the approximate eigenvalues hold \cite{SF73}:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}

\textcolor{red}{********* Hou voorlopig hier *************}
\subsubsection*{Properties of eigenvalues}
It is well known that the eigenvalues of a eigenvalue problem such as \eqref{GEVP} form an infinite sequence $(\lambda_i)$. Assume that the eigenvalues $\lambda_i$ can be ordered $\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...$ and denote the corresponding eigenfunctions $u_i$.

Since any multiple of a eigenfunction is still an eigenfunction, the eigenfunctions can be normalized so that $||u_i||_{W} = 1$ for all $i$.

@@

Also any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}
\textcolor{red}{********* Hou voorlopig hier *************}

\section{Estimating the eigenvalues.} \label{sec:estimating_the_eigenvalues}
In this section, the work in the textbook \cite{SF73} is discussed. The results are the same as given in the textbook, how ever the proofs are expanded for greater clarity.

\subsection{Projection of the eigenfunctions}

\subsubsection*{Rayleigh quotient}
\begin{equation}
	R(v) = \frac{b(v,v)}{c(v,v)} \ \text { for } v \in V. \label{Rayleigh}
\end{equation} \label{sym:Rayleigh}

\subsubsection*{Projection}
If $u \in V$, then $Pu$ is its projection in the subspace $S^h$.
\begin{equation*}
	b(u-Pu,v^h) = 0 \ \text{ for all } v \in S^h.
\end{equation*}

Let $E_j \in V$ denote the eigenspace spanned by the exact eigenvectors $\left\{u_1,u_2,...,u_j \right\}$ for $j = 1,2,...,m$. Clearly $m \leq N_e$.

Consider the subspace $S_j$ of $S^h$ where
\begin{equation*}
	S_j = PE_j \ \text{ for } j = 1,2,...,m.
\end{equation*}

The elements $Pu_j$ are the projections of the eigenfunctions $u_j$ into the space $S^h$. These projections $Pu_j$ are not necessarily equal to $u^h \in S^h$. In fact, $u^h_j$  can be vastly different from $u_j$. The situation is not simple. It is possible that $Py_k = 0$ for large $k$ Assume that the dimension of $S^h$ is large enough, substantially larger than $m$.

Let $B_m = \left\{u \in E_m \ | \ ||u||_{W} = 1 \right\}$ and define $\mu_m = \inf\left\{(Pu,Pu \ | \ u \in B_m)\right\}$.

The first step to obtain estimates for the eigenvalues, it to show that the elements of $B_m$ are linearly independent. In the first part we follow the approach in \cite{Zie2000}. The author introduced the quantity $\mu_m$ above.

\subsection{Upper bounds for approximate eigenvalues}

\newtheorem{Prop_2}[Prop_1]{Proposition}
\begin{Prop_2}
	\label{Prop_2}
	$\lambda^{h}_{m} \leq \max R(Pu) \ \text{ for } u \in B_{m}$
\end{Prop_2}
\begin{proof}
	Since $\dim S_{m} = m$, following from the minmax principle that
	\begin{equation}
		\lambda_m^h \leq \max R(v) \ \text{ for } v \in S_m. \label{eq:Prop_2}
	\end{equation}

	Now we take an arbitrary $v \in S_{m}$, $v \neq 0$. Then there exists a $u \in E_{m}$ such that $Py = v$. This $Py$ is the projection into $S_m$ of some $u \in E_m$ (which is also $\displaystyle \frac{1}{||u||_{W}}u \in B_m$).

	Next we show that $R(||u||_{W}^{-1}u) = R(u)$:

	\[ \frac{b\left(\frac{1}{||u||_{W}}u,\frac{1}{||u||_{W}}u\right)}{c\left(\frac{1}{||u||_{W}}u,\frac{1}{||u||_{W}}u\right)} = \frac{b(u,u)}{c(u,u)} = R(u). \]

	Finally, from \eqref{eq:Prop_2}
	\begin{eqnarray*}
	\lambda_{m}^{h} &\leq & \max R(v) \ \text{ for } \ v \in S_{m},\\
						&=& \max R(Pu) \ \text{ for } \ u \in E_{m},\\
						&=& \max R(Pu) \ \text{ for } \ u \in B_{m}.
	\end{eqnarray*}
\end{proof}

Pick any $i,j \in \mathbb{N}$ such that $i,j \leq m$. Then
\begin{eqnarray*}
	b(u_{i},\nu) &=& \lambda_{i}c(u_{i},\nu),\\
	\textrm{and } b(u_{j},\nu) &=& \lambda_{j}c(u_{j},\nu).
\end{eqnarray*}
for each $\nu \in V$. Clearly
\begin{eqnarray*}
	b(u_{i},u_{j}) &=& \lambda_{i}c(u_{i},u_{j}),\\
	\textrm{and } b(u_{j},u_{i}) &=& \lambda_{j}c(u_{j},u_{i}).
\end{eqnarray*}
Then using the symmetry of $b(\cdot,\cdot)$ and $c( \cdot, \cdot )$,
\begin{eqnarray*}
	0 &=& (\lambda_{i} - \lambda_{j}) c( u_{i}, u_{j} ).
\end{eqnarray*}
So if $i \neq j$ then $\lambda_{i} \neq \lambda_{j}$. Therefore $u_{i}$ and $u_{j}$ are orthogonal when $\lambda_i \neq \lambda_j$.

	The summation parameters can be merged into a single parameter. Then
	\begin{eqnarray*}
		b(u,u)  & = & \sum_{i=1}^{m} c_{i}^{2} \lambda_{i} u_i,\\
				& \leq & \lambda_{m}\sum_{i=1}^{m} c_{i}^{2} u_i,\\
				& = & \lambda_{m}||u||_{W}^2.
	\end{eqnarray*} for all $u \in B_m$.\\

	Using the Rayleigh quotient, and the definition of $\mu_{m}^{h}$, 
	\begin{eqnarray*}
		R(Pu) &=& \frac{b(Pu,Pu)}{c(Pu,Pu)},\\
			&=& \frac{b(Pu,Pu)}{||Pu||_{W}^{2}},\\
			&\leq & \frac{\lambda_{m}}{\mu_{m}^{h}}.
	\end{eqnarray*}

\subsection{The error bound}

Following Lemma \ref{Lem_1}, and since $\lambda_{i}^{h} \geq \lambda_{i}$ it follows that $0 < \mu_{m}^{h} \leq 1$. It is now possible to define the `error bound' in \cite{SF73}:
\begin{eqnarray}
	\sigma_{m}^{h} = 1 - \mu_{m}^{h}. \label{eq:error_bound}
\end{eqnarray}

\newtheorem{Cor_1}[Prop_1]{Proposition}
\begin{Cor_1}
	$0 \leq \sigma_{m}^{h} < 1$ and $\lambda_{m}^{h} - \lambda_{m} \leq \lambda_{m}^{h}\sigma_{m}^{h}$
\end{Cor_1}
\begin{proof}
	Starting with the result of Lemma \ref{Lem_1}, $\lambda_{m}^{h}\mu_{m}^{h} \leq \lambda_{m}$.

	Since $-\lambda_{m} \leq -\lambda_{m}^{h}\mu_{m}^{h}$, it follows that
	\begin{align*}
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h} - \lambda_{m}^{h}\mu_{m}^{h} = \lambda^h_m(1 - \mu^h_m). 
	\end{align*}
\end{proof}

\newtheorem{Prop_3}[Prop_1]{Proposition} 
\begin{Prop_3}
	\label{Prop_3}
	$\sigma_{m}^{h} = \max\left\{ 2c( u,u-Pu )-||u-Pu||_{W}^{2} \ | \ u \in B_{m} \right\}$
\end{Prop_3}
\begin{proof}
	Let $u \in B_{m}$. Then
	\begin{eqnarray*}
		||u - Pu||_{W}^{2} &=& c( u - Pu, u - Pu ), \\
						&=& c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ), \\
						&=& 2c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ) - c( u, u ),\\
						&=& 2c( u, u - Pu ) + c( Pu, Pu ) - c( u, u ).\\
	\end{eqnarray*}
	Consequently,
	\begin{eqnarray*}
		c( u, u ) - c( Pu, Pu )  & = & 2c( u, u - Pu ) - ||u - Pu||_{W}^{2}.
	\end{eqnarray*}

	Since $u \in B_{m}$, $||u||_{W}^{2} = 1$ and hence
	\begin{eqnarray*}
		1 - ||Pu||_{W}^{2}  & = & 2c( u, u - Pu ) - ||u - Pu||_{W}^{2}.
	\end{eqnarray*}

	On the right hand side, $1 - ||Pu||_{W}^{2} \leq 1 - \mu_{m}^{h} = \sigma_{m}^{h}$ for all $u \in B_{m}$. Therefore
	\begin{eqnarray*}
	\sigma_{m}^{h} = \max\left\{ 2c( u,u-Pu )-||u-Pu||_{W}^{2} \ | \ u \in B_{m} \right \}.
	\end{eqnarray*}
\end{proof}

Introduce some new notation for convenience. For any $u \in E_{m}$, let $u^{*} = \sum_{i=1}^{m} c_{i}\lambda_{i}^{-1}u_{i}$ where $u = \sum_{i=1}^{m} c_{i}u_{i}$.

\newtheorem{Lem_2}[Lem_1]{Lemma} 
\begin{Lem_2}
	\label{Lem_2}
	$\sigma_{m}^{h} \leq \max \left\{2||u^{*}- Pu^{*}||_{W} ||u-Pu||_{W} \ | \ u \in B_{m} \right\}.$
\end{Lem_2}
\begin{proof}
	From Proposition \ref{Prop_3},
	\begin{eqnarray*}
		\sigma_{m}^{h} & = & \max\left\{ 2c( u,u-Pu )-||u-Pu||_{W}^{2} \ | \ u \in B_{m} \right\}, \\
					& \leq & \max\left\{ 2c( u,u-Pu ) \ | \ u \in B_{m} \right\}. \\
	\end{eqnarray*}

	Using the Schwartz inequality,
	\begin{eqnarray*}
		b(u^{*} - Pu^{*}, u -Pu) & \leq & ||u^{*} - Pu^{*}||_{W}||u -Pu||_{W}.
	\end{eqnarray*}

	Finally,
	\begin{eqnarray*}
		\sigma_{m}^{h} & \leq & \max\left\{ 2||u^{*}- Pu^{*}||_{W} ||u-Pu||_{W} \ | \ u \in B_{m} \right\}.
	\end{eqnarray*}
\end{proof}

\subsection{Convergence of the eigenvalues}

\newtheorem{Prop_5}[Prop_1]{Proposition} 
\begin{Prop_5}
	\label{Prop_5}
	For any $\epsilon > 0$, there exists a $\delta >0$ such that for $h<\delta$,
	\begin{eqnarray*}
	||u^{*} - Pu^{*}||_{W} < \epsilon \ \textrm{ for each } u \in B_{m},\\
	||u - Pu||_{W} < \epsilon \ \textrm{ for each } u \in B_{m}.
	\end{eqnarray*}
\end{Prop_5}
\begin{proof}
	Consider a set of basis functions $\Phi \subset S^h$ such that $\phi^(k) = 0$ for each $\phi \in \Phi$.

	From the Rayleigh-Ritz Projection, $Pu$ is the closest element in $S^h$ to $u$. In particular, $||u-Pu||_{W}\leq ||u-\Pi u||_{W}$, where $\Pi u$ is the interpolation of $u$ into $S^h$.

	\textcolor{red}{Need to fix reference to approximation theorem. Need to wait until section 3.5 is done. The approximation theorem is moved to the end of section 3.5}

	From the Approximation Theorem given at the beginning of section \ref{sec:est}, is follows that
	\begin{eqnarray*}
		||u - \Pi u||_{W} \leq Ch^{k}||u^{(k)}||_{W}.
	\end{eqnarray*}

	So for any $\epsilon > 0$, a $\delta$ can be found such that if $h<\delta$ then
	\begin{eqnarray*}
		||u - Pu||_{W} \leq ||u - \Pi u||_{W} \leq \epsilon.
	\end{eqnarray*}
	A similar argument for $||u^{*} - Pu^{*}||_{W}$ proves the theorem.
\end{proof}

\newtheorem{Lem_3}[Lem_1]{Lemma} 
\begin{Lem_3}
	\label{Lem_3}
	For any $\epsilon >0$ there exists a $\delta > 0$ such that
	\begin{eqnarray}
	\sigma_{m}^{h} < \epsilon \ \textrm{ if } \ h < \delta. \label{eq:lem3}
	\end{eqnarray}
\end{Lem_3}
\begin{proof}
	For any $\epsilon > 0$ there exists a $\delta > 0$ such that if $h<\delta$, then
	\begin{eqnarray*}
	||u-Pu||_{W} < \epsilon \ \textrm{ for each } \ u \in B_{m}.
	\end{eqnarray*}
	This result, together with Lemma \ref{Lem_2} and Proposition \ref{Prop_5} gives the result.
\end{proof}

\newtheorem{Lem_4}[Lem_1]{Lemma} 
\begin{Lem_4}
	\label{Lem_4}
	There exists a $\delta > 0$ such that for $h < \delta$
	\begin{eqnarray}
	\lambda_{m}^{h} - \lambda_{m} \leq 2\lambda_{m}\sigma_{m}^{h}. \label{eq:lem4}
	\end{eqnarray}
\end{Lem_4}
\begin{proof}
	Using Lemma \ref{Lem_3}, choose $\delta$ such that $\sigma_{m}^{h} < \frac{1}{2}$. Then (by Lemma \ref{Lem_1}) $\lambda_{m}^{h} < 2\lambda_{m}$ and therefore $\lambda_{m}^{h} - \lambda_{m} \leq 2\lambda_{m}\sigma_{m}^{h}$.
\end{proof}

The convergence of the eigenvalues follows from \eqref{eq:lem3} and \eqref{eq:lem4}.

\section{Convergence of the eigenfunctions}
The next step is to show the convergence of the eigenfunctions. The problem can be formulated using the following result.

\newtheorem{Lem_5}[Lem_1]{Lemma} 
\begin{Lem_5}
	\label{Lem_5}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h} ) + \lambda_{m}^{h} - \lambda_{m}.
		\end{eqnarray*}
\end{Lem_5}
\begin{proof}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& b(u_{m},u_{m}) - 2b(u_{m},u^{h}_{m}) + b(u^{h}_{m},u^{h}_{m}), \\
										&=& \lambda_{m} c( u_{m}, u_{m} ) - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h}c( u_{m}^{h},u_{m}^{h} ),\\
										&=&  \lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h}, \\
										&=& 2\lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h} - \lambda_{m},\\
										&=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) + \lambda_{m}^{h} - \lambda_{m}.
		\end{eqnarray*}
\end{proof}

\newtheorem{Lem_6}[Lem_1]{Lemma} 
\begin{Lem_6}
	\label{Lem_6}
	For all m and j
	\begin{eqnarray*}
	(\lambda_{j}^{h} - \lambda_{m}) c( Pu_{m}, u_{j}^{h}) = \lambda_{m} c( u_{m}-Pu_{m},u_{j}^{h} ).
	\end{eqnarray*}
\end{Lem_6}
\begin{proof}
	Since the term $\lambda_{m}c( Pu,u^{h}_{j})$ appears on both sides of the equation, it is only required to show that
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& \lambda_{m} c( u, u_{j}^{h} ).	
	\end{eqnarray*}
	Since both $u$ and $u_{j}^{h}$ are eigenfunctions, then
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& b(Pu,u_{j}^{h}),\\
		\lambda_{m} c( u, u_{j}^{h} ) &=& b(u,u_{j}^{h}).
	\end{eqnarray*}

\newtheorem{Lem_7}[Lem_1]{Lemma} 
\begin{Lem_7}
	\label{Lem_7}
	\begin{eqnarray*}
		||Pu - \beta Pu^h_m||_{W}^{2} &\leq & {\rho}^{2} ||u_{m} - Pu_{m}||_{W}^{2}.
	\end{eqnarray*}
\end{Lem_7}
\begin{comment}
\begin{proof}

Using Lemma \ref{Lem_6}:
\begin{eqnarray*}
||Pu - \beta u_{m}^{h}||_{W}^{2} &=& \sum_{j\neq m} \left(\frac{\lambda_{m}}{|\lambda_{m} - \lambda_{j}^{h}|}\right)^{2} c( u_{m} - Pu_{m} ,u_{j}^{h})^{2},\\
				&\leq & \rho^{2} \sum_{j\neq m} c( u_{m} - Pu_{m} ,u_{j}^{h})^{2}, \\
				&\leq & \rho^{2} \sum_{j=1}^{N} c( u_{m} - Pu_{m} ,u_{j}^{h})^{2}, \\
				& = & \rho^{2} ||u_m - Pu_m||_{W}^{2}.\\
\end{eqnarray*}
\end{proof}
\end{comment}

\newtheorem{Lem_8}[Lem_1]{Lemma} 
\begin{Lem_8}
	\label{Lem_8}
	\begin{eqnarray*}
		||u_{m} - \beta u_{m}^{h}||_{W} &\leq & \left(1+\rho\right)||u_{m}-Pu_{m}||_{W}.
	\end{eqnarray*}
\end{Lem_8}
\begin{comment}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - \beta u_{m}^{h}||_{W} & \leq & ||u_{m}-Pu_{m}||_{W} + ||Pu_{m} - \beta u_{m}^{h}||_{W} \\
				& \leq & \left(1+\rho\right)||u_{m}-Pu_{m}||_{W} \ \ \textrm{ (by Lemma \ref{Lem_7})}
	\end{eqnarray*}
\end{proof}
\end{comment}

So again using the Approximation Theorem, it follows that $||u_{m} - \beta u_{m}^{h}||_{W}\leq Ch^{k}||u^{k}||_{W}$.

\newtheorem{Lem_9}[Lem_1]{Lemma} 
\begin{Lem_9}
	\label{Lem_9}
	\begin{eqnarray*}
		||u_{m} -  u_{m}^{h}||_{W} &\leq & 2||u_{m}-\beta u^{h}_{m}||_{W}.
	\end{eqnarray*}
\end{Lem_9}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - u_{m}^{h}||_{W} &=& ||u_{m} - \beta u_{m}^{h} + \beta u_{m}^{h} - u_{m}^{h}||_{W}, \\
						& \leq & ||u_{m} - \beta u_{m}^{h}||_{W} + ||\beta u_{m}^{h} - u_{m}^{h}||_{W}, \\
						& = & 2||u_{m} - \beta u_{m}^{h}||_{W}.	
	\end{eqnarray*}
\end{proof}

Therefore t $||u_{m} -  u_{m}^{h}||_{W} \leq Ch^{k}||u^{k}||_{W}$. So for any $\epsilon >0$ , a $\delta >0$ can be found such that if $h < \delta$, $||u_{m} -  u_{m}^{h}||_{W} < \epsilon$.

@@

For the requirements of this dissertation, this can be simplified with an assumption. The assumption is that the basis of $S^h$ consists of polynomials. With this assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||_{W}$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}||_{W} \leq C h^{k-m} || u^{(k)}||_{W} \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

This will be referred to as the approximation theorem in this section.
\end{document}

### 2023-11-26 files/Chapter5/2D_FEM.tex
Chapter 3

These functions are chosen as bi-cubic functions $\phi_i$. Simpler bi-linear functions can also be used, however the use of the bi-cubic basis functions results in faster convergence and also the benefit of obtaining the derivatives of the solution at the expense of more complexity.

@@

Consider a set of 
For this dissertation, we consider bi-cubic basis functions, however bi-linear basis functions can also be used.

>>>>

These functions are chosen as piecewise Hermite bi-cubic functions $\phi_i$. Simpler bi-linear functions can also be used, however the use of the bi-cubic basis functions results in faster convergence and also the benefit of obtaining the derivatives of the solution at the expense of more complexity.

@@

Consider a set of 
For this dissertation, we consider piecewise Hermite bi-cubic basis functions, however piecewise bi-linear basis functions can also be used.
